---
title: 'Clicking Analysis in Grain Boundary Search Task (E1)'
author: 'Dave Braun'
date: March 3, 2022
output:
  html_document:
    toc: true
    theme: flatly
    df_print: paged
    code_folding: hide
    includes:
      in_header: !expr paste0(here::here(), '/html/favicon.html')
      after_body: !expr paste0(here::here(), '/html/footer.html')
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(inputFile,
                encoding = encoding,
                output_file = 'index.html')})
---

This document was last updated on `r Sys.time()`.

In this document, we are exploring sequential effects within trials from clicking behavior. I'll look at RTs and error rates across line clicks, and try to assess things like whether they're clicking on the thinnest line first or not.


```{r include = FALSE}
library(tidyverse)
library(data.table)
library(ggridges)
library(ez)
```

```{r}
d <- read.csv('../../data/exp1_long_data.csv')
thinnest_lines <- read.csv('../../data/thinnest_lines.csv')
if (!'accuracy_type' %in% colnames(d)) {
  source('../compute_accuracy_columns.r')
  d <- compute_accuracy_columns(d, thinnest_lines)
  
}

N <- length(unique(d$participant))  
head(d)
head(thinnest_lines)
colnames(d)[which(colnames(d) == 'participant')] <- 'subject'
```

There were `r N` participants included in this analysis.

Only looking at trials where they selected only three lines:

```{r}
click_summary <- d %>% 
  group_by(subject, trial_count) %>% 
  summarize(count = n()) %>% 
  filter(count == 4) %>% 
  select(-count) %>% 
  mutate(key = 'keep')

d <- d %>% 
  inner_join(click_summary) %>% 
  filter(key == 'keep', is.na(accuracy)) %>% 
  select(-key)

```

## Accuracy and RT across clicks


```{r}

d$selection_rt_ms_relative <- ifelse(shift(d$click_order) < d$click_order, d$selection_rt_ms - shift(d$selection_rt_ms), d$selection_rt_ms)
d$selection_rt_ms_relative[1] <- d$selection_rt_ms[1]
d$accuracy_numeric <- ifelse(d$accuracy_type == 'hit', 1, 0)

subject_summary <- d %>% 
  group_by(subject, click_order) %>% 
  summarize(RT = mean(selection_rt_ms_relative), Accuracy = mean(accuracy_numeric)) 
subject_summary %>% 
  gather(outcome, value, RT:Accuracy) %>% 
  ggplot(aes(x = value, y = click_order, group = click_order)) + 
  geom_density_ridges(fill = 'steel blue', alpha = .8) + 
  facet_wrap(~outcome, scales = 'free') + 
  labs(
    x = '',
    y = 'Click Order'
  ) + 
  theme_bw() + 
  theme(strip.background = element_rect(color = 'black', fill = 'white'))

```


Some quick stats

```{r}
acc <- ezANOVA(wid = subject, within = click_order, dv = Accuracy, data = subject_summary)
rt <- ezANOVA(wid = subject, within = click_order, dv = RT, data = subject_summary)

acc
rt
```

Main effect of click order for both accuracy and RT.




## Click order and spatial effects

Looking at quadrants first

```{r}
source('../compute_quadrants.r')
line_data <- read.csv('../../data/exp1_line_data.csv')
d <- compute_quadrants(d, line_data)
```


```{r}
d %>% 
  group_by(subject, quadrant, click_order) %>% 
  summarize(count = n()) %>% 
  group_by(quadrant, click_order) %>% 
  summarize(average_sum = mean(count)) %>% 
  ggplot(aes(x = quadrant, y = average_sum, group = click_order)) + 
  geom_bar(stat = 'identity', aes(fill = factor(click_order)), color = 'black') + 
  scale_fill_brewer() + 
  theme_bw() + 
  labs(
    x = 'Area on Grid',
    y = 'Average Proportion',
    fill = 'Click Order'
  ) 
```


## Clicking and unclicking

Looking only at trials where they clicked and unclicked lines

```{r}
d <- read.csv('../../data/exp1_long_data.csv')
print(nrow(d))
trial_summary <- d %>% 
  group_by(trial_count, participant) %>% 
  summarize(count = n()) %>% 
  mutate(keep = ifelse(count > 4, 1, 0))

p_summary <- d %>% 
  filter(click_order == 1) %>% 
  group_by(participant) %>% 
  summarize(total_trials = n())

d <- trial_summary %>% 
  inner_join(d) %>% 
  filter(keep == 1) %>% 
  select(-keep)

t <- d %>% 
  filter(click_order == 1) %>% 
  group_by(participant) %>% 
  summarize(deselect_trials = n())

p_summary <- cbind(p_summary, t[,'deselect_trials'])

p_summary %>% 
  mutate(proportion_deselect = deselect_trials / total_trials) %>% 
  ggplot(aes(x = reorder(participant, proportion_deselect), y = proportion_deselect)) + 
  geom_bar(stat='identity') + 
  labs(
    x = 'Participant',
    y = 'Proportion of Trials Deselecting Lines'
  ) + 
  theme_bw() + 
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank())
```

Did people who deselected more often have faster click rts on average?

```{r}
p_summary <- d %>% 
  ## take out submit button presses
  filter(is.na(accuracy)) %>% 
  group_by(participant) %>% 
  summarize(selection_rt_ms = mean(selection_rt_ms)) %>% 
  inner_join(p_summary)

head(p_summary)
```

```{r}
p_summary %>% 
  mutate(proportion_deselect = deselect_trials / total_trials) %>% 
  ggplot(aes(x = proportion_deselect, y = selection_rt_ms)) + 
  geom_point(aes(group = participant)) + 
  labs(
    x = 'Proportion of Trials Deselecting Lines',
    y = 'Average Line Selection RT (ms)'
  ) + 
  theme_bw()
```
Oh wow pretty much the opposite. High deselecters look like they're maybe just deliberating more.

















