---
title: "Grain Boundary Search Task (V1) Timing Analysis"
author: "Dave Braun"
date: "12/13/2021"
output: 
  html_document:
    code_folding: hide
    df_print: paged
    theme: flatly
    includes:
      after_body: ../../../../../../html/footer.html
      in_header: ../../../../../../html/favicon.html
    
knit:
  (function(inputFile, encoding) {
      rmarkdown::render(inputFile,
                    encoding = encoding,
                    output_file = 'index.html')})
---

This document was last updated at `r Sys.time()`.

  
```{r include = FALSE}
library(tidyverse)
library(data.table)
```
  
```{r}
d <- read.csv('../../data/exp1_long_data.csv')
thinnest_lines <- read.csv('../../data/thinnest_lines.csv')
line_data <- read.csv('../../data/exp1_line_data.csv')
head(d)
N <- d %>% 
  group_by(participant) %>% 
  summarize(n()) %>% 
  nrow()
```

**There were `r N` participants in this study.**
  
```{r}
calc_accuracy_columns <- function(data) {
  participant <- data$participant[1]
  trial <- data$trial_count[1]
  
  thinnest_lines_set <- thinnest_lines[thinnest_lines$participant == participant & thinnest_lines$trial_count == trial,]$line_id
  
  ## determine the final three lines they selected
  ## some extra logic to assign NA to any lines that were selected and later released
  selected_lines <- c(NA, NA, NA)
  indices <- c(NA, NA, NA)
  for (row in 1:(nrow(data))) {
    if (data[row,]$selected_or_released == 'selected') {
      slot <- which(is.na(selected_lines))[1]
      selected_lines[slot]<- data[row,]$line_id
      indices[slot] <- row
    } else if (data[row,]$selected_or_released == 'released'){
      slot <- which(!(is.na(selected_lines)) & selected_lines == data[row,]$line_id)
      selected_lines[slot] <- NA
      indices[slot] <- NA
    }
  }
  
  selected_lines_accuracy_column <- rep(NA, nrow(data))
  accuracy_code <- ifelse(selected_lines %in% thinnest_lines_set, 'hit', 'false_alarm')
  selected_lines_accuracy_column[indices] <- accuracy_code

  return(selected_lines_accuracy_column)

}

out  <- vector()

for (participant in sapply(unique(d$participant), as.integer)) {
  for (trial in unique(d[d$participant == participant,]$trial_count)) {
    out <- c(out, calc_accuracy_columns(d[d$participant == participant & d$trial_count == trial,]))
  }
}

d$accuracy_type <- out


```
  

```{r}
## summarize hits and false alarms
summary_data <- d %>% 
  filter(!(is.na(accuracy_type))) %>% 
  mutate(hit = ifelse(accuracy_type == 'hit', 1, 0), false_alarm = ifelse(accuracy_type == 'false_alarm', 1, 0)) %>% 
  group_by(line_orientation) %>% 
  summarize(hits = sum(hit), false_alarms = sum(false_alarm))
```


**Coding other important variables.**

Did the trial include an indecision (ie, selecting and releasing)? A relative line RT measure.

```{r}
## calculate indecision trial
t <- d %>% 
  group_by(participant, trial_count, selected_or_released) %>% 
  summarize(count = n()) %>% 
  filter(selected_or_released == 'released') %>% 
  mutate(indecision_trial = 1) %>% 
  select(participant, trial_count, indecision_trial)

d <- d %>% 
  left_join(t)

d$indecision_trial <- ifelse(is.na(d$indecision_trial), 0, 1)
```

```{r}
## calculate relative line rt
d$selection_rt_ms_relative <- ifelse(d$trial_count == lag(d$trial_count), d$selection_rt_ms - lag(d$selection_rt_ms), d$selection_rt_ms)
```


## Selection RT over time

```{r}
d %>% 
  mutate(indecision_trial = factor(indecision_trial)) %>% 
  group_by(participant, trial_count, indecision_trial) %>% 
  summarize(selection_rt_ms = mean(selection_rt_ms_relative), prompt_rt = mean(prompt_rt_sec * 1000)) %>% 
  gather(phase, rt_ms, selection_rt_ms, prompt_rt) %>% 
  group_by(trial_count, indecision_trial, phase) %>% 
  summarize(rt = mean(rt_ms), se = sd(rt_ms) / sqrt(N)) %>% 
  ggplot(aes(x = trial_count, y = rt, group = indecision_trial)) + 
  geom_ribbon(aes(ymin = rt - se, ymax = rt + se, fill = indecision_trial), alpha = .5) + 
  geom_line(aes(color = indecision_trial)) + 
  facet_wrap(~phase, scales = 'free') + 
  labs(
    x = 'Trial',
    y = 'Response Time (ms)',
    fill = 'Indecision Trial',
    color = 'Indecision Trial'
  ) + 
  theme_bw() + 
  theme(legend.position = c(.8,.8),
        strip.background = element_rect(color = 'black', fill = 'white'))
  
    
  
```


Does time spent looking at the prompt negatively correlate with time selecting lines?

```{r}
d %>% 
  mutate(indecision_trial = factor(indecision_trial)) %>% #,
  mutate(indecision_trial = recode(indecision_trial, `0` = 'No Indecision', `1` = 'Indecision')) %>% 
  group_by(participant, trial_count, indecision_trial) %>% 
  summarize(selection_rt_ms = mean(selection_rt_ms_relative), prompt_rt_ms = mean(prompt_rt_sec * 1000), possible_thinnest_lines = max(possible_thinnest_lines)) %>% 
  group_by(trial_count, indecision_trial) %>% 
  summarize(selection_rt_ms = mean(selection_rt_ms), prompt_rt_ms = mean(prompt_rt_ms), possible_thinnest_lines = max(possible_thinnest_lines)) %>% 
  ggplot(aes(x = prompt_rt_ms, y = selection_rt_ms)) + 
  geom_point(aes(size = possible_thinnest_lines, color = possible_thinnest_lines)) + 
  geom_smooth(method = 'lm') + 
  facet_wrap(~indecision_trial) + 
  labs(
    x = 'Prompt RT (ms)',
    y = 'Selection RT (ms)',
    size = 'Possible Thinnest Lines',
    color = 'Possible Thinnest Lines',
    caption = 'Each dot represents a trial mean'
  ) + 
  theme_bw() + 
  theme(strip.background = element_rect(color = 'black', fill = 'white'),
        text = element_text(size = 20))
  
```
```{r}
ggsave('timing.pdf', height = 1000/96, width = 1920/2/96)
```



This makes me curious to look at frequencies for possible thinnest lines broken by indecision

```{r}
p <- d %>% 
  mutate(indecision_trial = factor(indecision_trial)) %>% 
  mutate(indecision_trial = recode(indecision_trial, `0` = 'No Indecision', `1` = 'Indecision')) %>% 
  group_by(participant, trial_count, indecision_trial) %>% 
  summarize(possible_thinnest_lines = max(possible_thinnest_lines)) %>% 
  group_by(indecision_trial, possible_thinnest_lines) %>% 
  summarize(count = n()) 

p2 <- p %>% 
  group_by(indecision_trial) %>% 
  summarize(overall_count = sum(count))

p %>% 
  inner_join(p2) %>% 
  mutate(proportion = count / overall_count) %>% 
  ggplot(aes(x = possible_thinnest_lines, y = proportion)) +
  geom_bar(stat = 'identity') + 
  facet_wrap(~indecision_trial) + 
  labs(
    x = 'Possible Thinnest Lines',
    y = 'Proportion of Trials'
  ) + 
  theme_bw() + 
  theme(strip.background = element_rect(color = 'black', fill = 'white'))
```

Yea wow these look exactly the same. okay.
















