---
title: Anomaly Analysis for Grain Boundary Visual Search (E3)
author: Dave Braun
date: November 12, 2021
output:
  html_document:
    theme: flatly
    code_folding: hide
    df_print: paged
    includes:
      after_body: ../../../../../html/footer.html
      in_header: ../../../../../html/favicon.html
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(inputFile,
              encoding = encoding,
              output_file = 'index.html')})
---

```{r include = FALSE}
library(tidyverse)
library(reticulate)
d <- read.csv('../../data/exp3_long_data.csv')
thinnest_lines <- read.csv('../../data/thinnest_lines.csv')
line_data <- read.csv('../../data/exp3_line_data.csv')
N <- length(unique(d$participant))
```


*This document was last updated at `r Sys.time()`.*

This is a space where we'll explore some of the effects related to the presence absence of the anomaly, which is simply just one inner hexagon boundary that's randomly chosen to be missing on 30% of trials. The lines that were selected to be among the anomalies were a subset of the total lines, mostly clustered in the middle of the array to minimize the number of lines that border the "larger grain" that also fell on the exterior of the whole array.  

**Sample size = `r N`.**

**Things to look at:**  

1. Start with just a simple analysis looking at whether there was a bias to be attracted to or averse to boundaries surrounding the larger grain relative to those other boundaries.


# 1. Simple anomaly analysis

Need to start with some of the basic preprocessing that's been done in the usual eda analyses, which involves summarizing which lines were selected on each trial.

```{r}
calc_accuracy_columns <- function(data) {
  participant <- data$participant[1]
  trial <- data$trial_count[1]
  
  thinnest_lines_set <- thinnest_lines[thinnest_lines$participant == participant & thinnest_lines$trial_count == trial,]$line_id
  
  ## determine the final three lines they selected
  ## some extra logic to assign NA to any lines that were selected and later released
  selected_lines <- c(NA, NA, NA)
  indices <- c(NA, NA, NA)
  for (row in 1:(nrow(data))) {
    if (data[row,]$selected_or_released == 'selected') {
      slot <- which(is.na(selected_lines))[1]
      selected_lines[slot]<- data[row,]$line_id
      indices[slot] <- row
    } else if (data[row,]$selected_or_released == 'released'){
      slot <- which(!(is.na(selected_lines)) & selected_lines == data[row,]$line_id)
      selected_lines[slot] <- NA
      indices[slot] <- NA
    }
  }
  
  selected_lines_accuracy_column <- rep(NA, nrow(data))
  accuracy_code <- ifelse(selected_lines %in% thinnest_lines_set, 'hit', 'false_alarm')
  selected_lines_accuracy_column[indices] <- accuracy_code

  return(selected_lines_accuracy_column)

}

out  <- vector()

for (participant in sapply(unique(d$participant), as.integer)) {
  for (trial in 0:99) {
    out <- c(out, calc_accuracy_columns(d[d$participant == participant & d$trial_count == trial,]))
  }
}

d$accuracy_type <- out

```


```{r}
compute_selected_lines <- function(data){
  participant <- data$participant[1]
  trial <- data$trial_count[1]

  selected_lines <- c(NA, NA, NA)
  indices <- c(NA, NA, NA)
  for (row in 1:(nrow(data))) {
    if (data[row,]$selected_or_released == 'selected') {
      slot <- which(is.na(selected_lines))[1]
      selected_lines[slot]<- data[row,]$line_id
      indices[slot] <- row
    } else if (data[row,]$selected_or_released == 'released'){
      slot <- which(!(is.na(selected_lines)) & selected_lines == data[row,]$line_id)
      selected_lines[slot] <- NA
      indices[slot] <- NA
    }
  }
  
  return(selected_lines)
  
}




trial_summary <- data.frame(participant = vector(), trial = vector(), is_anomaly_trial = vector(), anomaly_line_id = vector(), anomaly_group = vector(), selected_line_1 = vector(), selected_line_2 = vector(), selected_line_3 = vector())

for (participant in sapply(unique(d$participant), as.integer)) {
  for (trial in 0:99) {
    selected_lines <- compute_selected_lines(d[d$participant == participant & d$trial_count == trial,])
    is_anomaly_trial <- d[d$participant == participant & d$trial_count == trial,]$is_anomaly_trial[1]
    anomaly_line_id <- d[d$participant == participant & d$trial_count == trial,]$anomaly_line_id[1]
    anomaly_group <- d[d$participant == participant & d$trial_count == trial,]$anomaly_group[1]
    trial_summary <- rbind(trial_summary, data.frame(participant = participant, trial = trial, is_anomaly_trial = is_anomaly_trial, anomaly_line_id = anomaly_line_id, anomaly_group = anomaly_group,
                                                     selected_line_1 = selected_lines[1], selected_line_2 = selected_lines[2], selected_line_3 = selected_lines[3]))
  }
}

head(trial_summary)
```


```{python}
import pandas as pd
import pickle
import sys

trial_summary = pd.DataFrame(r.trial_summary)
trial_summary_dict = trial_summary.to_dict('records')

with open('contiguous_dict.pickle', 'rb') as file:
  contiguous_dict = pickle.load(file)



trial_summary_start = [x for x in trial_summary_dict if x['is_anomaly_trial'] == 'True']
trial_summary_end = []


for trial in trial_summary_start:
  num_contiguous = 0
  for i in range(1,4):
    if trial['selected_line_{}'.format(i)] in contiguous_dict[trial['anomaly_line_id']]:
      num_contiguous += 1
  trial['num_contiguous'] = num_contiguous
  trial_summary_end.append(trial)

trial_summary = pd.DataFrame(trial_summary_end)
```

```{r}
trial_summary_all <- trial_summary
trial_summary = py$trial_summary
trial_summary$num_non_contiguous <- 3 - trial_summary$num_contiguous
head(trial_summary)
```


```{r}
total_inner_lines <- nrow(line_data[line_data$participant == 59 & line_data$trial_count == 0 & line_data$is_exterior != 'True',])
total_trials <- nrow(trial_summary)
non_big_grain_lines_0 <- (total_inner_lines - 10 - 1) * total_trials
big_grain_lines_0 <- (10) * total_trials
non_big_grain_lines_1 <- (total_inner_lines - 9 - 1) * total_trials
big_grain_lines_1 <- 9 * total_trials
non_big_grain_lines_2 <- (total_inner_lines - 8 - 1) * total_trials
big_grain_lines_2 <- 8 * total_trials

results <- data.frame(expected = vector(), observed = vector(), group = vector())

for (group in 0:2){
  expected <- c(eval(as.name(paste0('non_big_grain_lines_', group))), eval(as.name(paste0('big_grain_lines_', group))))
  observed <- c(sum(trial_summary[trial_summary$anomaly_group == group,]$num_non_contiguous), sum(trial_summary[trial_summary$anomaly_group == group,]$num_contiguous))
  print(c(expected, observed))
  #results[paste0('test', group)] <- chisq.test(expected, observed)
  results <- rbind(results, data.frame(expected_non = expected[1], expected_anomaly = expected[2], observed_non = observed[1], observed_anomaly = observed[2], group = group))
}

results
```

**Overall count**

```{r}
results$expected_proportion <- results$expected_anomaly / (results$expected_non + results$expected_anomaly)
results$observed_proportion <- results$observed_anomaly / (results$observed_non + results$observed_anomaly)

plot_data <- results %>% 
  select(group, expected_proportion, observed_proportion) %>% 
  gather(var, value, expected_proportion, observed_proportion) 

plot_data %>% 
  mutate(var = recode(var, 'expected_proportion' = 'expected', 'observed_proportion' = 'observed')) %>% 
  ggplot(aes(x = group, y = value, group = var)) +
  geom_bar(stat = 'identity', aes(fill = var), position = position_dodge(width = .9)) + 
  ylim(0, .5) + 
  labs(
    x = 'Number of Borders of Anomaly Touching Exterior',
    y = 'Proportion of Anomaly Borders to \nall Selectable Borders',
    fill = ''
  ) + 
  theme_bw() + 
  theme(legend.position = 'bottom')
  

ggsave('basic_anomaly_result.pdf', height = 5.5, width = 7.5, units = 'in')
```

**Subject-level Proportions**

```{r}
library(ez)

expected_proportions <- data.frame(anomaly_group = c(0:2), expected_proportion = c(10 / 33, 9 / 33,  8 / 33))


anova_data <- trial_summary %>% 
  group_by(participant, anomaly_group) %>% 
  summarize(num_contiguous = sum(num_contiguous), num_non_contiguous = sum(num_non_contiguous)) %>% 
  mutate(observed_proportion = num_contiguous / (num_contiguous + num_non_contiguous)) %>% 
  inner_join(expected_proportions) %>% 
  gather(proportion_type, proportion, observed_proportion, expected_proportion)  

anova_data %>%     
  group_by(anomaly_group, proportion_type) %>% 
  summarize(mean_proportion = mean(proportion), se = sd(proportion) / sqrt(length(unique(trial_summary$participant)))) %>% 
  ggplot(aes(x = anomaly_group, y = mean_proportion)) + 
  geom_bar(stat = 'identity', aes(fill = proportion_type), position = position_dodge(width = .9)) + 
  geom_errorbar(aes(ymin = mean_proportion - se, ymax = mean_proportion + se, group = proportion_type), position = position_dodge(width = .9), width = .5) + 
  ylim(0, .5) + 
  labs(
    x = 'Anomaly Group',
    y = 'Mean Proportion of Anomaly\n Borders to All Available Boundaries',
    fill = 'Proportion Type'
  ) + 
  theme_bw()

  


```


**one sample t tests**

```{r include = FALSE}
ttest_data <- anova_data %>% 
  spread(proportion_type, proportion)

ttest_results <- list()

for (group in 0:2) {
  mu <- ttest_data[ttest_data$anomaly_group == group,]$expected_proportion[1]
  ttest_results[group + 1] <- t.test(x = ttest_data[ttest_data$anomaly_group == group,]$observed_proportion, mu = mu)$p.value
}

ttest_results
```


The observed proportion for anomaly group 1 is significantly different from what we'd expect based on chance.


## Anomalies and biases

Doing a basic check to see whether the anomaly groups are disproportionately distributed along the quadrants or line orientations.

```{python}
import pickle

with open('contiguous_dict.pickle', 'rb') as file:
  contiguous_dict = pickle.load(file)

d1 = []
d2 = []

for key in contiguous_dict:
  d1 += [key] * 10
  d2 += contiguous_dict[key]
  
contiguous = pd.DataFrame({'anomaly_lines': d1, 'contiguous_lines': d2})
```

```{r}
## merge in group number
contiguous <- py$contiguous
group_summary <- trial_summary %>% 
  group_by(anomaly_line_id) %>% 
  summarize(anomaly_group = unique(anomaly_group))
colnames(group_summary)[1] <- 'anomaly_lines'
contiguous <- contiguous %>% 
  inner_join(group_summary)
head(contiguous)
```

```{r}
## merge in exterior information
exterior_summary <- line_data %>% 
  mutate(center_x = (top_x + bottom_x) / 2, center_y = (top_y + bottom_y) / 2) %>% 
  group_by(line_id) %>% 
  summarize(is_exterior = unique(is_exterior), center_x = unique(center_x), center_y = unique(center_y), line_orientation = unique(line_orientation))
colnames(exterior_summary)[1] <- 'contiguous_lines'
contiguous <- contiguous %>% 
  inner_join(exterior_summary)
head(contiguous)
```

```{r}
## get quadrants
top_left <- c(min(line_data$top_x), max(line_data$top_y))
top_right <- c(max(line_data$top_x), max(line_data$top_y))
bottom_left <- c(min(line_data$bottom_x), min(line_data$bottom_y))
bottom_right <- c(max(line_data$bottom_x), min(line_data$bottom_y))
center <- c(mean(c(top_left[1], top_right[1])), mean(c(top_left[2], bottom_left[2])))

contiguous <- contiguous %>% 
  mutate(quadrant = ifelse(center_x > center[1] & center_y > center[2], 'I', ifelse(center_x < center[1] & center_y > center[2], 'II',
                    ifelse(center_x < center[1] & center_y < center[2], 'III', ifelse(center_x > center[1] & center_y < center[2], 'IV', '')))),
         xdist = abs(center_x - center[1]), 
         ydist = abs(center_y - center[2])) %>% 
  mutate(is_on_axis = ifelse(xdist < 0.1, 'y_axis', ifelse(ydist < 0.1, 'x_axis', ''))) %>% 
  mutate(quadrant = ifelse(is_on_axis == '', quadrant, is_on_axis)) %>% 
  select(-center_x, -center_y, xdist, ydist) 
```

```{r}
## line orientation
contiguous %>% 
  filter(is_exterior == 'False') %>% 
  group_by(anomaly_group, line_orientation) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x = anomaly_group, y = count)) + 
  geom_bar(stat = 'identity', aes(fill = line_orientation), position = position_dodge(width = .9)) +
  labs(
    x = 'Anomaly Group',
    y = 'Count of Lines',
    fill = 'Line Orientation'
  ) + 
  theme_bw()
```


```{r}
## quadrant
contiguous %>% 
  filter(is_exterior == 'False') %>% 
  group_by(anomaly_group, quadrant) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x = anomaly_group, y = count)) + 
  geom_bar(stat = 'identity', aes(fill = quadrant), position = position_dodge(width = .9)) +
  labs(
    x = 'Anomaly Group',
    y = 'Count of Lines',
    fill = 'Quadrant'
  ) + 
  theme_bw()
```















