---
title: 'Correlation Analysis in Grain Boundary Search Task (E1)'
author: 'Dave Braun'
date: April 28, 2022
output:
  html_document:
    toc: true
    theme: flatly
    df_print: paged
    code_folding: hide
    includes:
      in_header: !expr paste0(here::here(), '/html/favicon.html')
      after_body: !expr paste0(here::here(), '/html/footer.html')
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(inputFile,
                encoding = encoding,
                output_file = 'index.html')})
---

This document was last updated on `r Sys.time()`.

Here I'll be taking several different metrics from the GB search task and correlating them to assess the relationship between different behaviors in the task. A few things to look at:  
  * Are differences in response times (in line clicking) associated with selection biases (eg, over attending to x-axis)?  
    * I can look at average rt at the subject level and also subjects grouped by their rt cluster groups (which preserves temporal information).  
  * Are selection biases associated with accuracies?
  
```{r include = FALSE}
library(tidyverse)
```


```{r}
d <- read.csv('../../data/exp1_long_data.csv')
line <- read.csv('../../data/exp1_line_data.csv')
head(d)
```


```{r}
## calculate quadrants
source('../compute_quadrants.r')
d <- compute_quadrants(d, line)
colnames(d)
```


## Response time and selection bias

```{r}
## to compute individual selection bias
  ## calculate variance between observed and expected.. right?
  ## as var approaches zero choices are unbiased across the grid, as var increases choices are more biased
```

























