---
title: 'Clustering Analysis in Grain Boundary Search Task (E1)'
author: 'Dave Braun'
date: March 28, 2022
output:
  html_document:
    toc: true
    theme: flatly
    df_print: paged
    code_folding: hide
    includes:
      in_header: !expr paste0(here::here(), '/html/favicon.html')
      after_body: !expr paste0(here::here(), '/html/footer.html')
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(inputFile,
                encoding = encoding,
                output_file = 'index.html')})
---

This document was last updated on `r Sys.time()`.

In this document, we are exploring clustering subjects together based on various features:  
  
  * Using hierarchical clustering to compare subjects on line clicking RTs across trials
  
# Hierarchical Clustering

**Some considerations**  
I'm going to start by only thinking about response times over time, although I could imagine adding in accuracy as a next step. I'm going to bring in the raw data, because there's a need to have rts from all trials for all subjects to make an $N_{Subjects}~X~T_{Trials}$ matrix. This means I'll need to handle trials with excessively long rts- I'm thinking I'll impute the subject's median for those trials. I'll also only consider the first three line clicks in a trial.


```{r include = FALSE}
library(tidyverse)
library(data.table)
library(grid)
library(ggdendro)
```

```{r}
d <- read.csv('../../data/exp1_long_data_raw.csv')
thinnest_lines <- read.csv('../../data/thinnest_lines.csv')
if (!'accuracy_type' %in% colnames(d)) {
  source('../compute_accuracy_columns.r')
  out <- compute_accuracy_columns(d, thinnest_lines)
  
}
d <- out[[1]]
thinnest_lines <- out[[2]]
N <- length(unique(d$participant))  
head(d)
head(thinnest_lines)
colnames(d)[which(colnames(d) == 'participant')] <- 'subject'
colnames(d)[which(colnames(d) == 'selection_rt_ms')] <- 'selection_rt_ms_absolute'
d$selection_rt_ms <- ifelse(shift(d$trial_count) == d$trial_count, d$selection_rt_ms_absolute - shift(d$selection_rt_ms_absolute), d$selection_rt_ms_absolute)
d$selection_rt_ms[1] <- d$selection_rt_ms_absolute[1]
```

There were `r N` participants included in this analysis.


## Keep only first three line selections

```{r}
d$prompt_rt_ms <- d$prompt_rt_sec * 1000
```


```{r}

keep_three_selections <- function(d) {
  indices <- c()
  for (i in 1:(nrow(d))) {
    if (d[i,]$selected_or_released == 'selected') {
      indices <- c(indices, i)
    }
    if (length(indices) == 3) {
      break
    }
  }
  return(d[indices,])
}


sub_d <- select(d, subject, trial_count, selection_rt_ms, prompt_rt_ms, selected_or_released)
new_d <- data.frame(subject = c(), trial_count = c(), prompt_rt_ms = c(), selection_rt_ms = c())

for (subject in unique(sub_d$subject)) {
  for (trial in unique(sub_d[sub_d$subject == subject,]$trial_count)) {
    new_d <- rbind(new_d, keep_three_selections(sub_d[sub_d$subject == subject & d$trial_count == trial,]))
  }
}
```

## Handle outliers

Doing a median imputation for outside 2 sds on group means

```{r}
selection_rt_ms_avg <- mean(new_d$selection_rt_ms)
selection_rt_ms_sd <- sd(new_d$selection_rt_ms)
prompt_rt_ms_avg <- mean(new_d$prompt_rt_ms)
prompt_rt_ms_sd <- sd(new_d$prompt_rt_ms)

new_d <- new_d %>% 
  group_by(subject) %>% 
  summarize(selection_median = median(selection_rt_ms), prompt_median = median(prompt_rt_ms)) %>% 
  inner_join(new_d) %>% 
  rename(selection_rt_ms_raw = selection_rt_ms, prompt_rt_ms_raw = prompt_rt_ms) %>% 
  mutate(selection_rt_ms = ifelse(selection_rt_ms_raw >= selection_rt_ms_avg + 2*selection_rt_ms_sd | selection_rt_ms_raw <= selection_rt_ms_avg - 2*selection_rt_ms_sd, selection_median, selection_rt_ms_raw),
         prompt_rt_ms = ifelse(prompt_rt_ms_raw >= prompt_rt_ms_avg + 2*prompt_rt_ms_sd | prompt_rt_ms_raw <= prompt_rt_ms_avg - 2*prompt_rt_ms_sd, prompt_median, prompt_rt_ms_raw)) %>% 
  select(-selection_rt_ms_raw, -prompt_rt_ms_raw, -selection_median, -prompt_median)
```


## Create matrix for clustering

```{r}
click_levels = c()
for (i in 0:99) {
  click_levels = c(click_levels, paste0('Trial', i, '_ClickOne'), paste0('Trial', i, '_ClickTwo'), paste0('Trial', i, '_ClickThree'))
}


prompt_levels <- paste0('Trial', 0:99, 'PromptRTMS')
```

```{r}
t <- new_d %>% 
  mutate(trial_char = paste0('Trial', new_d$trial_count)) %>% 
  select(-selected_or_released, -prompt_rt_ms, -trial_count) %>% 
  mutate(click_count = rep(c('ClickOne', 'ClickTwo', 'ClickThree'), nrow(new_d) / 3)) %>% 
  unite(spreadable, trial_char, click_count, sep='_') %>% 
  mutate(spreadable = factor(spreadable, levels = click_levels)) %>% 
  spread(spreadable, selection_rt_ms) 
  
t2 <- new_d %>% 
  group_by(subject, trial_count) %>% 
  summarize(prompt_rt_ms = unique(prompt_rt_ms)) %>% 
  mutate(trial_char = paste0('Trial', trial_count, 'PromptRTMS')) %>% 
  select(-trial_count) %>% 
  mutate(trial_char = factor(trial_char, levels = prompt_levels)) %>% 
  spread(trial_char, prompt_rt_ms) 
  

cluster_d <- data.frame(cbind(t2, t[,colnames(t) != 'subject']))
rownames(cluster_d) <- cluster_d$subject
cluster_d <- cluster_d[,colnames(cluster_d) != 'subject']
cluster_d <- scale(cluster_d)
```


## Execute clustering

```{r}
## take only t data
t <- data.frame(t)
rownames(t) <- t$subject
t <- t[,colnames(t) != 'subject']
hc <- as.dendrogram(hclust(dist(t)))

#hc <- as.dendrogram(hclust(dist(cluster_d)))
dendro_plot <- ggdendrogram(data = hc, rotate = TRUE) + 
  theme(axis.text.x = element_blank())
print(dendro_plot)
```

Hmm I need more than this.

Resourcing [this](https://jcoliver.github.io/learn-r/008-ggplot-dendrograms-and-heatmaps.html)

```{r}
subject_order <- order.dendrogram(hc)


heatmap <- data.frame(cluster_d) %>% 
  mutate(subject = factor(rownames(cluster_d), levels = rownames(cluster_d)[subject_order], ordered = TRUE)) %>% 
  gather(measure, value, Trial0PromptRTMS:Trial99_ClickThree) %>% 
  ggplot(aes(x = measure, y = subject)) + 
  geom_tile(aes(fill = value)) + 
  scale_fill_gradient2() + 
  labs(
    x = 'Time in Experiment',
    y = 'Subject',
    fill = 'Standarized Line Click RT'
  ) + 
  theme_bw() + 
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        legend.position = 'bottom')
```

```{r}
library(gridExtra)
grid.arrange(heatmap, dendro_plot, nrow = 1)
```


```{r}
grid.newpage()
print(heatmap, 
      vp = viewport(x = 0.4, y = 0.5, width = 0.8, height = 1.0))
print(dendro_plot, 
      vp = viewport(x = 0.90, y = .53, width = 0.2, height = 1))

```


















