---
title: "GB Search Task (E1) OPAM Analyses"
author: "Dave Braun"
date: "08/10/2022"
output: 
  html_document:
    code_folding: hide
    df_print: paged
    theme: flatly
    includes:
      after_body: ../../../../../../html/footer.html
      in_header: ../../../../../../html/favicon.html
    
knit:
  (function(inputFile, encoding) {
      rmarkdown::render(inputFile,
                    encoding = encoding,
                    output_file = 'index.html')})
---

This document was last updated at `r Sys.time()`.

The big effects of interest:  

  * Quadrant bias  
  * Line orientation bias
  
```{r include = FALSE}
library(tidyverse)
library(testit)
library(psycho)
```

  
  
```{r}
d <- read.csv('../../data/exp1_long_data.csv')
thinnest_lines <- read.csv('../../data/thinnest_lines.csv')
line_data <- read.csv('../../data/exp1_line_data.csv')
head(d)
N <- length(unique(d$participant))
```

**There were `r N` participants in this study.**
  
```{r}
source('../compute_accuracy_columns.r')

result <- compute_accuracy_columns(d, thinnest_lines)
d <- result[[1]]
thinnest_lines <- result[[2]]

```
  
```{r}
## a function for spot checking accuracy coding throughout the data
spot_check <- function() {
  participant <- sample(d$participant, size = 1)
  trial <- sample(d[d$participant==participant,]$trial_count, size = 1)
  d_sub <- d[d$participant == participant & d$trial_count == trial,]
  t_sub <- thinnest_lines[thinnest_lines$participant == participant & thinnest_lines$trial_count == trial,]
  print('Subset of long data:')
  print(d_sub[,c('line_id', 'selected_or_released', 'accuracy_type')])
  print(' ')
  print(' ')
  print('Subset of thinnest lines data:')
  print(t_sub[,c('line_width', 'line_id', 'accuracy_type')])
}
## Anything coded as a "hit" in the long data should be represented in the thinnest lines data
## Anything coded as a "false alarm" shouldn't be represented in the thinnest lines data
spot_check()
```


```{r}
## summarize hits and false alarms
summary_data <- d %>% 
  filter(!(is.na(accuracy_type))) %>% 
  mutate(hit = ifelse(accuracy_type == 'hit', 1, 0), false_alarm = ifelse(accuracy_type == 'false_alarm', 1, 0)) %>% 
  group_by(participant, line_orientation) %>% 
  summarize(hits = sum(hit), false_alarms = sum(false_alarm)) 

```

Base rate of line orientation types on each trial:

```{r}
line_data %>% 
  filter(participant == 10, trial_count == 2) %>% 
  group_by(line_orientation) %>% 
  summarize(count = n()) 
```


  
```{r}
## summarize misses
summary_data_p <- thinnest_lines %>% 
  filter(accuracy_type == 'miss') %>% 
  group_by(participant, line_orientation) %>% 
  summarize(miss_count = n()) %>% 
  inner_join(summary_data) %>% 
  gather(accuracy_type, count, miss_count:false_alarms) 
summary_data <- summary_data_p %>% 
  group_by(line_orientation, accuracy_type) %>% 
  summarize(m = mean(count), se = sd(count) / sqrt(N))

## export in long format so they can work with it for the poster
write.csv(summary_data_p, 'line_orientation.csv', row.names = FALSE)


summary_data %>% 
  ggplot(aes(x = line_orientation, y = m)) +
  geom_bar(stat = 'identity', aes(fill = accuracy_type), position = position_dodge(width = .9)) +
  geom_errorbar(aes(ymin = m - se, ymax = m + se, group = accuracy_type), position = position_dodge(width = .9), width = .5) + 
  labs(
    y = 'Mean Frequency',
    x = 'Line Orientation',
    fill = 'Accuracy Type'
  ) + 
  theme_bw() + 
  theme(legend.position = 'bottom')
  
```
  
```{r}
ggsave('line_orientation.png', height = 1080 / 300, width = 1920 / 300, units = 'in', dpi = 300)
```
  
~ *left off here* ~
  
  
## Quadrant Analysis

  
We'll calculate the center of the grid by looking at the most extreme X's and Y's and taking the center, then compare the center of the lines against that center point.
  
  
```{r}
line_data <- read.csv('../../data/exp1_line_data.csv')
head(line_data)
```

  
```{r}
top_left <- c(min(line_data$top_x), max(line_data$top_y))
top_right <- c(max(line_data$top_x), max(line_data$top_y))
bottom_left <- c(min(line_data$bottom_x), min(line_data$bottom_y))
bottom_right <- c(max(line_data$bottom_x), min(line_data$bottom_y))
print(top_left)
print(top_right)
print(bottom_left)
print(bottom_right)
```

```{r}
center <- c(mean(c(top_left[1], top_right[1])), mean(c(top_left[2], bottom_left[2])))
center
```


```{r}
source('../compute_quadrants.r')
## calculate line centers

d <- d %>% 
  mutate(line_center_x = (top_x + bottom_x) / 2, line_center_y = (top_y + bottom_y) / 2) %>% 
  mutate(quadrant = ifelse(line_center_x > center[1] & line_center_y > center[2], 'I', ifelse(line_center_x < center[1] & line_center_y > center[2], 'II',
                    ifelse(line_center_x < center[1] & line_center_y < center[2], 'III', ifelse(line_center_x > center[1] & line_center_y < center[2], 'IV', '')))))
line_data <- line_data %>% 
  mutate(line_center_x = (top_x + bottom_x) / 2, line_center_y = (top_y + bottom_y) / 2) %>% 
  mutate(quadrant = ifelse(line_center_x > center[1] & line_center_y > center[2], 'I', ifelse(line_center_x < center[1] & line_center_y > center[2], 'II',
                    ifelse(line_center_x < center[1] & line_center_y < center[2], 'III', ifelse(line_center_x > center[1] & line_center_y < center[2], 'IV', '')))))

## append line coordinates to thinnest lines data (for including misses in summarization)
## ~~~ LEFT OFF HERE ~~~

line_data %>% 
  filter(participant == 1, trial_count == 0, is_exterior == 'False') %>% 
  mutate(x_dist = abs(line_center_x - center[1]), y_dist = abs(line_center_y - center[2])) %>% 
  mutate(is_on_axis = ifelse(x_dist < 0.1, 'y_axis', ifelse(y_dist < 0.1, 'x_axis', ''))) %>% 
  mutate(quadrant = ifelse(is_on_axis == '', quadrant, is_on_axis)) %>% 
  group_by(quadrant) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x = quadrant, y = count)) + 
  geom_bar(stat = 'identity') + 
  labs(
    x = 'Quadrant',
    y = 'Number of Lines per Trial'
  ) + 
  theme_bw()

ggsave('lines_per_quadrant.png', height = 1080 / 300, width = 1920 / 300, units = 'in', dpi = 300)

d <- d %>% 
  mutate(x_dist = abs(line_center_x - center[1]), y_dist = abs(line_center_y - center[2])) %>% 
  mutate(is_on_axis = ifelse(x_dist < 0.1, 'y_axis', ifelse(y_dist < 0.1, 'x_axis', ''))) %>% 
  mutate(quadrant = ifelse(is_on_axis == '', quadrant, is_on_axis)) 
  
d %>%   
  filter(!(is.na(quadrant))) %>% 
  group_by(participant, quadrant) %>% 
  summarize(count = n()) %>% 
  group_by(quadrant) %>% 
  summarize(mean_trials = mean(count)) %>% 
  ggplot(aes(x = quadrant, y = mean_trials)) +
  geom_bar(stat = 'identity') +
  labs(
    x = 'Quadrant',
    y = 'Mean Frequency Selected across Participants'
  ) + 
  theme_bw()

ggsave('selection_by_quadrant.png', height = 1080 / 300, width = 1920 / 300, units = 'in', dpi = 300)

d %>% 
  filter(!(is.na(quadrant)), !(is.na(top_x)), !(is.na(accuracy_type))) %>% 
  group_by(participant, quadrant, accuracy_type) %>% 
  summarize(count = n()) %>% 
  group_by(quadrant, accuracy_type) %>% 
  summarize(mean_count = mean(count), se_count = sd(count) / sqrt(N)) %>% 
  ggplot(aes(x = quadrant, y = mean_count, group = accuracy_type)) +
  geom_bar(stat = 'identity', aes(fill = accuracy_type), position = position_dodge(width = .9)) + 
  geom_errorbar(aes(ymin = mean_count - se_count, ymax = mean_count + se_count, group = accuracy_type), width = .5, position = position_dodge(.9)) + 
  labs(
    x = 'Quadrant',
    y = 'Average Across Participants',
    fill = 'Accuracy Type'
  ) + 
  theme_bw() + 
  theme(legend.position = 'bottom')

ggsave('accuracy_by_quadrant.png', height = 1080 / 300, width = 1920 / 300, units = 'in', dpi = 300)

```
  

## Features of Possible Thinnest Lines
  
```{r}
thinnest_lines_summaries <- thinnest_lines %>% 
  group_by(participant, trial_count) %>% 
  summarize(mean_line_width = mean(line_width), sd_line_width = sd(line_width))
head(thinnest_lines_summaries)

```



```{r}

plot_data <- d %>% 
  group_by(participant, trial_count, accuracy_type) %>% 
  summarize(count = n(), possible_thinnest_lines = max(possible_thinnest_lines)) %>% 
  filter(!(is.na(accuracy_type))) %>% 
  inner_join(thinnest_lines_summaries) 

plot_data %>% 
  gather(variable, value, possible_thinnest_lines:sd_line_width) %>% 
  ggplot(aes(x = value)) +
  geom_histogram() + 
  facet_wrap(~variable, scales = 'free') + 
  labs(
    x = '',
    y = 'Frequency'
  ) +
  theme_bw() +
  theme(strip.background = element_rect(color = 'black', fill = 'white'))


```
```{r}
plot_data %>% 
  #filter(possible_thinnest_lines < 4) %>% 
  group_by(accuracy_type, count) %>% 
  summarize(frequency = n()) %>% 
  ggplot(aes(x = count, y = frequency)) + 
  geom_bar(stat = 'identity', aes(fill = accuracy_type), position = position_dodge(width = .9)) + 
  labs(
    x = 'Number of occurances in a trial',
    y = 'Frequency',
    fill = 'Accuracy Type'
  ) + 
  theme_bw()
```

```{r}
plot_data <- plot_data %>% 
  spread(accuracy_type, count) %>% 
  mutate(false_alarm = replace_na(false_alarm, 0),
         hit = replace_na(hit, 0),
         miss = ifelse(hit == 3, 0, possible_thinnest_lines - hit),
         correct_reject = (33 - false_alarm) - possible_thinnest_lines) %>% 
  mutate(d_prime = dprime(n_hit = hit, n_fa = false_alarm, n_miss = miss, n_cr = correct_reject)$dprime) 
```


```{r}
## average mean line width and d' to participant level and correlate

plot_data %>% 
  gather(var, value, false_alarm, correct_reject, possible_thinnest_lines, d_prime) %>% 
  ggplot(aes(x = value)) + 
  geom_histogram() + 
  facet_wrap(~var, scales = 'free') +
  theme_bw() + 
  theme(strip.background = element_rect(color = 'black', fill = 'white'))
  
plot_data %>% 
  gather(var, value, mean_line_width, sd_line_width, possible_thinnest_lines) %>% 
  ggplot(aes(x = value, y = d_prime)) +
  geom_point() +
  geom_smooth(method = 'lm') + 
  facet_wrap(~var, scales = 'free') +
  labs(
    x = '',
    y = "d'",
    caption = 'Each point represents a trial.'
  ) + 
  theme_bw() + 
  theme(strip.background = element_rect(color = 'black', fill = 'white'))

  
```

Try to look at the relationship between mean line width and d' while controlling for possible thinnest lines.

```{r}
plot_data %>% 
  mutate(possible_thinnest_lines_bin = ifelse(possible_thinnest_lines == 3, 'three', ifelse(possible_thinnest_lines == 4, 'four', ifelse(possible_thinnest_lines > 5 & possible_thinnest_lines < 7, 'five and six', '> six')))) %>% 
  mutate(possible_thinnest_lines_bin = factor(possible_thinnest_lines_bin, levels = c('three', 'four', 'five and six', '> six'))) %>% 
  ggplot(aes(x = mean_line_width, y = d_prime)) + 
  geom_point() + 
  geom_smooth(method = 'lm') + 
  facet_wrap(~possible_thinnest_lines_bin) + 
  labs(
    x = 'Mean line width',
    y = "d'",
    caption = 'Facets represent groupings on number of possible thinnest lines in a trial'
  ) + 
  theme_bw() + 
  theme(strip.background = element_rect(color = 'black', fill = 'white'))
```

  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  