---
title: "Grain Boundary Search Task (V1) EDA"
author: "Dave Braun"
date: "10/21/2021"
output: 
  html_document:
    code_folding: hide
    df_print: paged
    theme: flatly
    includes:
      after_body: ../../../../../html/footer.html
      in_header: ../../../../../html/favicon.html
    
knit:
  (function(inputFile, encoding) {
      rmarkdown::render(inputFile,
                    encoding = encoding,
                    output_file = 'index.html')})
---

This document was last updated at `r Sys.time()`.

## What do we want to look at?

* Why did Nick's participants take so much longer than Lily's?  
  * Somewhat captured in the plots already produced  
* IVs:  
  * Possible line widths  
  * Location of lines  
  * Line orientation
* DVs:  
  * Accuracy  
  * RT  
  * Choice  
  * Order of line selection  
* How does accuracy change over line selections and de selections

## Line orientation
This is not simple.

What we need:  
* The final three lines they selected  
* The widths and orientations of all possible thinnest lines they could've selected  
* With those two sets, calculate hits, misses, false alarms
  
```{r include = FALSE}
library(tidyverse)
library(testit)
library(psycho)
```

  
  
```{r}
d <- read.csv('../../data/exp1_long_data.csv')
## THINNEST LINES IS BROKEN
thinnest_lines <- read.csv('../../data/thinnest_lines.csv')
line_data <- read.csv('../../data/exp1_line_data.csv')
head(d)
N <- d %>% 
  group_by(participant) %>% 
  summarize(n()) %>% 
  nrow()
```

**There were `r N` participants in this study.**
  
```{r}
source('../compute_accuracy_columns.r')

d <- compute_accuracy_columns(d, thinnest_lines)

```
  
```{r}
## a function for spot checking accuracy coding throughout the data
spot_check <- function() {
  participant <- sample(d$participant, size = 1)
  trial <- sample(d[d$participant==participant,]$trial_count, size = 1)
  d_sub <- d[d$participant == participant & d$trial_count == trial,]
  t_sub <- thinnest_lines[thinnest_lines$participant == participant & thinnest_lines$trial_count == trial,]
  print(d_sub[,c('line_id', 'selected_or_released', 'accuracy_type')])
  print(' ')
  print(' ')
  print(t_sub[,c('line_width', 'line_id')])
}
```


```{r}
## summarize hits and false alarms
summary_data <- d %>% 
  filter(!(is.na(accuracy_type))) %>% 
  mutate(hit = ifelse(accuracy_type == 'hit', 1, 0), false_alarm = ifelse(accuracy_type == 'false_alarm', 1, 0)) %>% 
  group_by(line_orientation) %>% 
  summarize(hits = sum(hit), false_alarms = sum(false_alarm))
```

Base rate of line orientation types on each trial:

```{r}
line_data %>% 
  filter(participant == 10, trial_count == 0) %>% 
  group_by(line_orientation) %>% 
  summarize(count = n()) 
```


  
```{r}
## summarize misses

thinnest_lines <- d %>% 
  filter(pressed_object == 'submit') %>% 
  select(participant, trial_count, accuracy) %>% 
  inner_join(thinnest_lines)

summary_data <- thinnest_lines %>% 
  filter(accuracy != 1) %>% 
  group_by(line_orientation) %>% 
  summarize(misses = n()) %>% 
  inner_join(summary_data) %>% 
  gather(accuracy_type, count, misses:false_alarms) 


d %>% 
  filter(selected_or_released == 'selected') %>% 
  group_by(participant, line_orientation) %>% 
  summarize(count = n()) %>% 
  group_by(line_orientation) %>% 
  summarize(mean_selection_rate = mean(count), se = sd(count) / sqrt(length(unique(d$participant)))) %>% 
  ggplot(aes(x = line_orientation, y = mean_selection_rate)) + 
  geom_bar(stat = 'identity') + 
  geom_errorbar(aes(ymin = mean_selection_rate - se, ymax = mean_selection_rate + se), width = .5) +
  labs(
    x = 'Line Orientation',
    y = 'Mean Selection Rate'
  ) + 
  theme_bw()

summary_data %>% 
  ggplot(aes(x = line_orientation, y = count)) +
  geom_bar(stat = 'identity', aes(fill = accuracy_type), position = position_dodge(width = .9)) +
  labs(
    y = 'Count',
    x = 'Line Orientation',
    fill = 'Accuracy Type'
  ) + 
  theme_bw() + 
  theme(legend.position = 'bottom')
  
```
  
```{r}
ggsave('line_orientation.png', height = 1080 / 300, width = 1920 / 300, units = 'in', dpi = 300)
```
  
  
  
  
## Quadrant Analysis

  
We'll calculate the center of the grid by looking at the most extreme X's and Y's and taking the center, then compare the center of the lines against that center point.
  
  
```{r}
line_data <- read.csv('../../data/exp1_line_data.csv')
head(line_data)
```

  
```{r}
top_left <- c(min(line_data$top_x), max(line_data$top_y))
top_right <- c(max(line_data$top_x), max(line_data$top_y))
bottom_left <- c(min(line_data$bottom_x), min(line_data$bottom_y))
bottom_right <- c(max(line_data$bottom_x), min(line_data$bottom_y))
print(top_left)
print(top_right)
print(bottom_left)
print(bottom_right)
```

```{r}
center <- c(mean(c(top_left[1], top_right[1])), mean(c(top_left[2], bottom_left[2])))
center
```


```{r}
## calculate line centers

d <- d %>% 
  mutate(line_center_x = (top_x + bottom_x) / 2, line_center_y = (top_y + bottom_y) / 2) %>% 
  mutate(quadrant = ifelse(line_center_x > center[1] & line_center_y > center[2], 'I', ifelse(line_center_x < center[1] & line_center_y > center[2], 'II',
                    ifelse(line_center_x < center[1] & line_center_y < center[2], 'III', ifelse(line_center_x > center[1] & line_center_y < center[2], 'IV', '')))))
line_data <- line_data %>% 
  mutate(line_center_x = (top_x + bottom_x) / 2, line_center_y = (top_y + bottom_y) / 2) %>% 
  mutate(quadrant = ifelse(line_center_x > center[1] & line_center_y > center[2], 'I', ifelse(line_center_x < center[1] & line_center_y > center[2], 'II',
                    ifelse(line_center_x < center[1] & line_center_y < center[2], 'III', ifelse(line_center_x > center[1] & line_center_y < center[2], 'IV', '')))))


line_data %>% 
  filter(participant == 1, trial_count == 0, is_exterior == 'False') %>% 
  mutate(x_dist = abs(line_center_x - center[1]), y_dist = abs(line_center_y - center[2])) %>% 
  mutate(is_on_axis = ifelse(x_dist < 0.1, 'y_axis', ifelse(y_dist < 0.1, 'x_axis', ''))) %>% 
  mutate(quadrant = ifelse(is_on_axis == '', quadrant, is_on_axis)) %>% 
  group_by(quadrant) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x = quadrant, y = count)) + 
  geom_bar(stat = 'identity') + 
  labs(
    x = 'Quadrant',
    y = 'Number of Lines per Trial'
  ) + 
  theme_bw()

ggsave('lines_per_quadrant.png', height = 1080 / 300, width = 1920 / 300, units = 'in', dpi = 300)

d <- d %>% 
  mutate(x_dist = abs(line_center_x - center[1]), y_dist = abs(line_center_y - center[2])) %>% 
  mutate(is_on_axis = ifelse(x_dist < 0.1, 'y_axis', ifelse(y_dist < 0.1, 'x_axis', ''))) %>% 
  mutate(quadrant = ifelse(is_on_axis == '', quadrant, is_on_axis)) 
  
d %>%   
  filter(!(is.na(quadrant))) %>% 
  group_by(participant, quadrant) %>% 
  summarize(count = n()) %>% 
  group_by(quadrant) %>% 
  summarize(mean_trials = mean(count)) %>% 
  ggplot(aes(x = quadrant, y = mean_trials)) +
  geom_bar(stat = 'identity') +
  labs(
    x = 'Quadrant',
    y = 'Mean Frequency Selected across Participants'
  ) + 
  theme_bw()

ggsave('selection_by_quadrant.png', height = 1080 / 300, width = 1920 / 300, units = 'in', dpi = 300)

d %>% 
  filter(!(is.na(quadrant)), !(is.na(top_x)), !(is.na(accuracy_type))) %>% 
  group_by(participant, quadrant, accuracy_type) %>% 
  summarize(count = n()) %>% 
  group_by(quadrant, accuracy_type) %>% 
  summarize(mean_count = mean(count), se_count = sd(count) / sqrt(N)) %>% 
  ggplot(aes(x = quadrant, y = mean_count, group = accuracy_type)) +
  geom_bar(stat = 'identity', aes(fill = accuracy_type), position = position_dodge(width = .9)) + 
  geom_errorbar(aes(ymin = mean_count - se_count, ymax = mean_count + se_count, group = accuracy_type), width = .5, position = position_dodge(.9)) + 
  labs(
    x = 'Quadrant',
    y = 'Average Across Participants',
    fill = 'Accuracy Type'
  ) + 
  theme_bw() + 
  theme(legend.position = 'bottom')

ggsave('accuracy_by_quadrant.png', height = 1080 / 300, width = 1920 / 300, units = 'in', dpi = 300)

```
  

## Features of Possible Thinnest Lines
  
```{r}
thinnest_lines_summaries <- thinnest_lines %>% 
  group_by(participant, trial_count) %>% 
  summarize(mean_line_width = mean(line_width), sd_line_width = sd(line_width))
head(thinnest_lines_summaries)

```



```{r}

plot_data <- d %>% 
  group_by(participant, trial_count, accuracy_type) %>% 
  summarize(count = n(), possible_thinnest_lines = max(possible_thinnest_lines)) %>% 
  filter(!(is.na(accuracy_type))) %>% 
  inner_join(thinnest_lines_summaries) 

plot_data %>% 
  gather(variable, value, possible_thinnest_lines:sd_line_width) %>% 
  ggplot(aes(x = value)) +
  geom_histogram() + 
  facet_wrap(~variable, scales = 'free') + 
  labs(
    x = '',
    y = 'Frequency'
  ) +
  theme_bw() +
  theme(strip.background = element_rect(color = 'black', fill = 'white'))


```
```{r}
plot_data %>% 
  #filter(possible_thinnest_lines < 4) %>% 
  group_by(accuracy_type, count) %>% 
  summarize(frequency = n()) %>% 
  ggplot(aes(x = count, y = frequency)) + 
  geom_bar(stat = 'identity', aes(fill = accuracy_type), position = position_dodge(width = .9)) + 
  labs(
    x = 'Number of occurances in a trial',
    y = 'Frequency',
    fill = 'Accuracy Type'
  ) + 
  theme_bw()
```

```{r}
plot_data <- plot_data %>% 
  spread(accuracy_type, count) %>% 
  mutate(false_alarm = replace_na(false_alarm, 0),
         hit = replace_na(hit, 0),
         miss = ifelse(hit == 3, 0, possible_thinnest_lines - hit),
         correct_reject = (33 - false_alarm) - possible_thinnest_lines) %>% 
  mutate(d_prime = dprime(n_hit = hit, n_fa = false_alarm, n_miss = miss, n_cr = correct_reject)$dprime) 
```


```{r}
## average mean line width and d' to participant level and correlate

plot_data %>% 
  gather(var, value, false_alarm, correct_reject, possible_thinnest_lines, d_prime) %>% 
  ggplot(aes(x = value)) + 
  geom_histogram() + 
  facet_wrap(~var, scales = 'free') +
  theme_bw() + 
  theme(strip.background = element_rect(color = 'black', fill = 'white'))
  
plot_data %>% 
  gather(var, value, mean_line_width, sd_line_width, possible_thinnest_lines) %>% 
  ggplot(aes(x = value, y = d_prime)) +
  geom_point() +
  geom_smooth(method = 'lm') + 
  facet_wrap(~var, scales = 'free') +
  labs(
    x = '',
    y = "d'",
    caption = 'Each point represents a trial.'
  ) + 
  theme_bw() + 
  theme(strip.background = element_rect(color = 'black', fill = 'white'))

  
```

Try to look at the relationship between mean line width and d' while controlling for possible thinnest lines.

```{r}
plot_data %>% 
  mutate(possible_thinnest_lines_bin = ifelse(possible_thinnest_lines == 3, 'three', ifelse(possible_thinnest_lines == 4, 'four', ifelse(possible_thinnest_lines > 5 & possible_thinnest_lines < 7, 'five and six', '> six')))) %>% 
  mutate(possible_thinnest_lines_bin = factor(possible_thinnest_lines_bin, levels = c('three', 'four', 'five and six', '> six'))) %>% 
  ggplot(aes(x = mean_line_width, y = d_prime)) + 
  geom_point() + 
  geom_smooth(method = 'lm') + 
  facet_wrap(~possible_thinnest_lines_bin) + 
  labs(
    x = 'Mean line width',
    y = "d'",
    caption = 'Facets represent groupings on number of possible thinnest lines in a trial'
  ) + 
  theme_bw() + 
  theme(strip.background = element_rect(color = 'black', fill = 'white'))
```

  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  