---
title: "Grain Boundary Search Task (V1) SSM Analysis"
author: "Dave Braun"
date: "10/21/2021"
output: 
  html_document:
    code_folding: hide
    df_print: paged
    theme: flatly
    includes:
      after_body: ../../../../../../html/footer.html
      in_header: ../../../../../../html/favicon.html
    
knit:
  (function(inputFile, encoding) {
      rmarkdown::render(inputFile,
                    encoding = encoding,
                    output_file = 'index.html')})
---

This document was last updated at `r Sys.time()`.

The goal here is to analyze these data for the presence of subsequent search misses (SSM; see Adamo et al., 2021). I would think that this effect would show up in our data as accuracy decreasing as a function of selection order of the lines.
  
```{r include = FALSE}
library(tidyverse)
library(testit)
library(psycho)
```

  
  
```{r}
d <- read.csv('../../data/exp1_long_data.csv')
thinnest_lines <- read.csv('../../data/thinnest_lines.csv')
line_data <- read.csv('../../data/exp1_line_data.csv')
head(d)
N <- d %>% 
  group_by(participant) %>% 
  summarize(n()) %>% 
  nrow()
```

**There were `r N` participants in this study.**
  
```{r}
calc_accuracy_columns <- function(data) {
  participant <- data$participant[1]
  trial <- data$trial_count[1]
  
  thinnest_lines_set <- thinnest_lines[thinnest_lines$participant == participant & thinnest_lines$trial_count == trial,]$line_id
  
  ## determine the final three lines they selected
  ## some extra logic to assign NA to any lines that were selected and later released
  selected_lines <- c(NA, NA, NA)
  indices <- c(NA, NA, NA)
  for (row in 1:(nrow(data))) {
    if (is.na(data[row,]$selected_or_released == 'selected')){
      print(c(participant, trial))
    }
    if (data[row,]$selected_or_released == 'selected') {
      slot <- which(is.na(selected_lines))[1]
      selected_lines[slot]<- data[row,]$line_id
      indices[slot] <- row
    } else if (data[row,]$selected_or_released == 'released'){
      slot <- which(!(is.na(selected_lines)) & selected_lines == data[row,]$line_id)
      selected_lines[slot] <- NA
      indices[slot] <- NA
    }
    
  }
  
  selected_lines_accuracy_column <- rep(NA, nrow(data))
  accuracy_code <- ifelse(selected_lines %in% thinnest_lines_set, 'hit', 'false_alarm')
  selected_lines_accuracy_column[indices] <- accuracy_code

  return(selected_lines_accuracy_column)

}

out  <- vector()

for (participant in sapply(unique(d$participant), as.integer)) {
  for (trial in sapply(unique(d[d$participant == participant,]$trial_count), as.integer)) {
    out <- c(out, calc_accuracy_columns(d[d$participant == participant & d$trial_count == trial,]))
  }
}

d$accuracy_type <- out


```
  

```{r}
## summarize hits and false alarms
summary_data <- d %>% 
  filter(!(is.na(accuracy_type))) %>% 
  mutate(hit = ifelse(accuracy_type == 'hit', 1, 0), false_alarm = ifelse(accuracy_type == 'false_alarm', 1, 0)) %>% 
  group_by(line_orientation) %>% 
  summarize(hits = sum(hit), false_alarms = sum(false_alarm))
```

Base rate of line orientation types on each trial:

```{r}
line_data %>% 
  filter(participant == 10, trial_count == 0) %>% 
  group_by(line_orientation) %>% 
  summarize(count = n()) 
```


  
```{r}
## summarize misses

thinnest_lines <- d %>% 
  filter(pressed_object == 'submit') %>% 
  select(participant, trial_count, accuracy) %>% 
  inner_join(thinnest_lines)

summary_data <- thinnest_lines %>% 
  filter(accuracy != 1) %>% 
  group_by(line_orientation) %>% 
  summarize(misses = n()) %>% 
  inner_join(summary_data) %>% 
  gather(accuracy_type, count, misses:false_alarms) 


d %>% 
  filter(selected_or_released == 'selected') %>% 
  group_by(participant, line_orientation) %>% 
  summarize(count = n()) %>% 
  group_by(line_orientation) %>% 
  summarize(mean_selection_rate = mean(count), se = sd(count) / sqrt(length(unique(d$participant)))) %>% 
  ggplot(aes(x = line_orientation, y = mean_selection_rate)) + 
  geom_bar(stat = 'identity') + 
  geom_errorbar(aes(ymin = mean_selection_rate - se, ymax = mean_selection_rate + se), width = .5) +
  labs(
    x = 'Line Orientation',
    y = 'Mean Selection Rate'
  ) + 
  theme_bw()

summary_data %>% 
  ggplot(aes(x = line_orientation, y = count)) +
  geom_bar(stat = 'identity', aes(fill = accuracy_type), position = position_dodge(width = .9)) +
  labs(
    y = 'Count',
    x = 'Line Orientation',
    fill = 'Accuracy Type'
  ) + 
  theme_bw() + 
  theme(legend.position = 'bottom')
  
```
  
```{r}
ggsave('line_orientation.png', height = 1080 / 300, width = 1920 / 300, units = 'in', dpi = 300)
```
  
  
  
  
## Quadrant Analysis

  
We'll calculate the center of the grid by looking at the most extreme X's and Y's and taking the center, then compare the center of the lines against that center point.
  
  
```{r}
line_data <- read.csv('../../data/exp1_line_data.csv')
head(line_data)
```

  
```{r}
top_left <- c(min(line_data$top_x), max(line_data$top_y))
top_right <- c(max(line_data$top_x), max(line_data$top_y))
bottom_left <- c(min(line_data$bottom_x), min(line_data$bottom_y))
bottom_right <- c(max(line_data$bottom_x), min(line_data$bottom_y))
print(top_left)
print(top_right)
print(bottom_left)
print(bottom_right)
```

```{r}
center <- c(mean(c(top_left[1], top_right[1])), mean(c(top_left[2], bottom_left[2])))
center
```


```{r}
## calculate line centers

d <- d %>% 
  mutate(line_center_x = (top_x + bottom_x) / 2, line_center_y = (top_y + bottom_y) / 2) %>% 
  mutate(quadrant = ifelse(line_center_x > center[1] & line_center_y > center[2], 'I', ifelse(line_center_x < center[1] & line_center_y > center[2], 'II',
                    ifelse(line_center_x < center[1] & line_center_y < center[2], 'III', ifelse(line_center_x > center[1] & line_center_y < center[2], 'IV', '')))))
line_data <- line_data %>% 
  mutate(line_center_x = (top_x + bottom_x) / 2, line_center_y = (top_y + bottom_y) / 2) %>% 
  mutate(quadrant = ifelse(line_center_x > center[1] & line_center_y > center[2], 'I', ifelse(line_center_x < center[1] & line_center_y > center[2], 'II',
                    ifelse(line_center_x < center[1] & line_center_y < center[2], 'III', ifelse(line_center_x > center[1] & line_center_y < center[2], 'IV', '')))))


line_data %>% 
  filter(participant == 1, trial_count == 0, is_exterior == 'False') %>% 
  mutate(x_dist = abs(line_center_x - center[1]), y_dist = abs(line_center_y - center[2])) %>% 
  mutate(is_on_axis = ifelse(x_dist < 0.1, 'y_axis', ifelse(y_dist < 0.1, 'x_axis', ''))) %>% 
  mutate(quadrant = ifelse(is_on_axis == '', quadrant, is_on_axis)) %>% 
  group_by(quadrant) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x = quadrant, y = count)) + 
  geom_bar(stat = 'identity') + 
  labs(
    x = 'Quadrant',
    y = 'Number of Lines per Trial'
  ) + 
  theme_bw()

ggsave('lines_per_quadrant.png', height = 1080 / 300, width = 1920 / 300, units = 'in', dpi = 300)

d <- d %>% 
  mutate(x_dist = abs(line_center_x - center[1]), y_dist = abs(line_center_y - center[2])) %>% 
  mutate(is_on_axis = ifelse(x_dist < 0.1, 'y_axis', ifelse(y_dist < 0.1, 'x_axis', ''))) %>% 
  mutate(quadrant = ifelse(is_on_axis == '', quadrant, is_on_axis)) 
  
d %>%   
  filter(!(is.na(quadrant))) %>% 
  group_by(participant, quadrant) %>% 
  summarize(count = n()) %>% 
  group_by(quadrant) %>% 
  summarize(mean_trials = mean(count)) %>% 
  ggplot(aes(x = quadrant, y = mean_trials)) +
  geom_bar(stat = 'identity') +
  labs(
    x = 'Quadrant',
    y = 'Mean Frequency Selected across Participants'
  ) + 
  theme_bw()

ggsave('selection_by_quadrant.png', height = 1080 / 300, width = 1920 / 300, units = 'in', dpi = 300)

d %>% 
  filter(!(is.na(quadrant)), !(is.na(top_x)), !(is.na(accuracy_type))) %>% 
  group_by(participant, quadrant, accuracy_type) %>% 
  summarize(count = n()) %>% 
  group_by(quadrant, accuracy_type) %>% 
  summarize(mean_count = mean(count), se_count = sd(count) / sqrt(N)) %>% 
  ggplot(aes(x = quadrant, y = mean_count, group = accuracy_type)) +
  geom_bar(stat = 'identity', aes(fill = accuracy_type), position = position_dodge(width = .9)) + 
  geom_errorbar(aes(ymin = mean_count - se_count, ymax = mean_count + se_count, group = accuracy_type), width = .5, position = position_dodge(.9)) + 
  labs(
    x = 'Quadrant',
    y = 'Average Across Participants',
    fill = 'Accuracy Type'
  ) + 
  theme_bw() + 
  theme(legend.position = 'bottom')

ggsave('accuracy_by_quadrant.png', height = 1080 / 300, width = 1920 / 300, units = 'in', dpi = 300)

```
  

## Features of Possible Thinnest Lines
  
```{r}
thinnest_lines_summaries <- thinnest_lines %>% 
  group_by(participant, trial_count) %>% 
  summarize(mean_line_width = mean(line_width), sd_line_width = sd(line_width))
head(thinnest_lines_summaries)

```



```{r}

plot_data <- d %>% 
  group_by(participant, trial_count, accuracy_type) %>% 
  summarize(count = n(), possible_thinnest_lines = max(possible_thinnest_lines)) %>% 
  filter(!(is.na(accuracy_type))) %>% 
  inner_join(thinnest_lines_summaries) 

plot_data %>% 
  gather(variable, value, possible_thinnest_lines:sd_line_width) %>% 
  ggplot(aes(x = value)) +
  geom_histogram() + 
  facet_wrap(~variable, scales = 'free') + 
  labs(
    x = '',
    y = 'Frequency'
  ) +
  theme_bw() +
  theme(strip.background = element_rect(color = 'black', fill = 'white'))


```
```{r}
plot_data %>% 
  #filter(possible_thinnest_lines < 4) %>% 
  group_by(accuracy_type, count) %>% 
  summarize(frequency = n()) %>% 
  ggplot(aes(x = count, y = frequency)) + 
  geom_bar(stat = 'identity', aes(fill = accuracy_type), position = position_dodge(width = .9)) + 
  labs(
    x = 'Number of occurances in a trial',
    y = 'Frequency',
    fill = 'Accuracy Type'
  ) + 
  theme_bw()
```

```{r}
plot_data <- plot_data %>% 
  spread(accuracy_type, count) %>% 
  mutate(false_alarm = replace_na(false_alarm, 0),
         hit = replace_na(hit, 0),
         miss = ifelse(hit == 3, 0, possible_thinnest_lines - hit),
         correct_reject = (33 - false_alarm) - possible_thinnest_lines) %>% 
  mutate(d_prime = dprime(n_hit = hit, n_fa = false_alarm, n_miss = miss, n_cr = correct_reject)$dprime) 
```


```{r}
## average mean line width and d' to participant level and correlate

plot_data %>% 
  gather(var, value, false_alarm, correct_reject, possible_thinnest_lines, d_prime) %>% 
  ggplot(aes(x = value)) + 
  geom_histogram() + 
  facet_wrap(~var, scales = 'free') +
  theme_bw() + 
  theme(strip.background = element_rect(color = 'black', fill = 'white'))
  
plot_data %>% 
  gather(var, value, mean_line_width, sd_line_width, possible_thinnest_lines) %>% 
  ggplot(aes(x = value, y = d_prime)) +
  geom_point() +
  geom_smooth(method = 'lm') + 
  facet_wrap(~var, scales = 'free') +
  labs(
    x = '',
    y = "d'",
    caption = 'Each point represents a trial.'
  ) + 
  theme_bw() + 
  theme(strip.background = element_rect(color = 'black', fill = 'white'))

  
```

Try to look at the relationship between mean line width and d' while controlling for possible thinnest lines.

```{r}
plot_data %>% 
  mutate(possible_thinnest_lines_bin = ifelse(possible_thinnest_lines == 3, 'three', ifelse(possible_thinnest_lines == 4, 'four', ifelse(possible_thinnest_lines > 5 & possible_thinnest_lines < 7, 'five and six', '> six')))) %>% 
  mutate(possible_thinnest_lines_bin = factor(possible_thinnest_lines_bin, levels = c('three', 'four', 'five and six', '> six'))) %>% 
  ggplot(aes(x = mean_line_width, y = d_prime)) + 
  geom_point() + 
  geom_smooth(method = 'lm') + 
  facet_wrap(~possible_thinnest_lines_bin) + 
  labs(
    x = 'Mean line width',
    y = "d'",
    caption = 'Facets represent groupings on number of possible thinnest lines in a trial'
  ) + 
  theme_bw() + 
  theme(strip.background = element_rect(color = 'black', fill = 'white'))
```

  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  