---
title: "Grain Boundary Search Task (V1) EDA"
author: "Dave Braun"
date: "9/24/2021"
output: 
  html_document:
    code_folding: hide
    df_print: paged
    theme: flatly
    
knit:
  (function(inputFile, encoding) {
      rmarkdown::render(inputFile,
                    encoding = encoding,
                    output_file = 'index.html')})
---


## What do we want to look at?

* Why did Nick's participants take so much longer than Lily's?  
  * Somewhat captured in the plots already produced  
* IVs:  
  * Possible line widths  
  * Location of lines  
  * Line orientation
* DVs:  
  * Accuracy  
  * RT  
  * Choice  
  * Order of line selection  
* How does accuracy change over line selections and de selections

## Line orientation
This is not simple.

What we need:  
* The final three lines they selected  
* The widths and orientations of all possible thinnest lines they could've selected  
* With those two sets, calculate hits, misses, false alarms
  
```{r include = FALSE}
library(tidyverse)
library(testit)
```

  
  
```{r}
d <- read.csv('../../data/raw/long_data/pilot/pilot_data.csv')
thinnest_lines <- read.csv('../../data/raw/line_data/pilot/thinnest_lines.csv')
head(d)
N <- d %>% 
  group_by(participant) %>% 
  summarize(n()) %>% 
  nrow()
```

  
```{r}
calc_accuracy_columns <- function(d) {
  participant <- d$participant[1]
  trial <- d$trial_count[1]
  
  thinnest_lines_set <- thinnest_lines[thinnest_lines$participant == participant & thinnest_lines$trial_count == trial,]$line_id
  
  ## determine the final three lines they selected
  ## some extra logic to assign NA to any lines that were selected and later released
  selected_lines <- c(NA, NA, NA)
  indices <- c(NA, NA, NA)
  for (row in 1:(nrow(d))) {
    if (d[row,]$selected_or_released == 'selected') {
      slot <- which(is.na(selected_lines))[1]
      selected_lines[slot]<- d[row,]$line_id
      indices[slot] <- row
    } else {
      slot <- which(!(is.na(selected_lines)) & selected_lines == d[row,]$line_id)
      selected_lines[slot] <- NA
      indices[slot] <- NA
    }
  }
  
  selected_lines_accuracy_column <- rep(NA, nrow(d))
  accuracy_code <- ifelse(selected_lines %in% thinnest_lines_set, 'hit', 'false_alarm')
  selected_lines_accuracy_column[indices] <- accuracy_code

  return(selected_lines_accuracy_column)

}

out  <- vector()

for (participant in unlist(lapply(levels(factor(d$participant)), as.integer))) {
  for (trial in 0:99) {
    out <- c(out, calc_accuracy_columns(d[d$participant == participant & d$trial_count == trial,]))
  }
}

d$accuracy_type <- out


```
  

```{r}
## summarize hits and false alarms
summary_data <- d %>% 
  filter(!(is.na(accuracy_type))) %>% 
  mutate(hit = ifelse(accuracy_type == 'hit', 1, 0), false_alarm = ifelse(accuracy_type == 'false_alarm', 1, 0)) %>% 
  group_by(line_orientation) %>% 
  summarize(hits = sum(hit), false_alarms = sum(false_alarm))
```



  
```{r}
## summarize misses

thinnest_lines <- d %>% 
  filter(pressed_object == 'submit') %>% 
  select(participant, trial_count, accuracy) %>% 
  inner_join(thinnest_lines)

summary_data <- thinnest_lines %>% 
  filter(accuracy != 1) %>% 
  group_by(line_orientation) %>% 
  summarize(misses = n()) %>% 
  inner_join(summary_data) %>% 
  gather(accuracy_type, count, misses:false_alarms) 

summary_data %>% 
  ggplot(aes(x = line_orientation, y = count)) +
  geom_bar(stat = 'identity', aes(fill = accuracy_type), position = position_dodge(width = .9)) +
  labs(
    y = 'Count',
    x = 'Line Orientation',
    fill = 'Accuracy Type'
  ) + 
  theme_bw() + 
  theme(legend.position = 'bottom')
  
```
  
```{r}
ggsave('line_orientation.png', height = 1080 / 300, width = 1920 / 300, units = 'in', dpi = 300)
```
  
  
  
  
## Quadrant Analysis

  
We'll calculate the center of the grid by looking at the most extreme X's and Y's and taking the center, then compare the center of the lines against that center point.
  
  
```{r}
line_data <- read.csv('../../data/raw/line_data/pilot/pilot_data.csv')
head(line_data)
```

  
```{r}
top_left <- c(min(line_data$top_x), max(line_data$top_y))
top_right <- c(max(line_data$top_x), max(line_data$top_y))
bottom_left <- c(min(line_data$bottom_x), min(line_data$bottom_y))
bottom_right <- c(max(line_data$bottom_x), min(line_data$bottom_y))
print(top_left)
print(top_right)
print(bottom_left)
print(bottom_right)
```

```{r}
center <- c(mean(c(top_left[1], top_right[1])), mean(c(top_left[2], bottom_left[2])))
center
```


```{r}
## calculate line centers

d <- d %>% 
  mutate(line_center_x = (top_x + bottom_x) / 2, line_center_y = (top_y + bottom_y) / 2) %>% 
  mutate(quadrant = ifelse(line_center_x > center[1] & line_center_y > center[2], 'I', ifelse(line_center_x < center[1] & line_center_y > center[2], 'II',
                    ifelse(line_center_x < center[1] & line_center_y < center[2], 'III', ifelse(line_center_x > center[1] & line_center_y < center[2], 'IV', '')))))
line_data <- line_data %>% 
  mutate(line_center_x = (top_x + bottom_x) / 2, line_center_y = (top_y + bottom_y) / 2) %>% 
  mutate(quadrant = ifelse(line_center_x > center[1] & line_center_y > center[2], 'I', ifelse(line_center_x < center[1] & line_center_y > center[2], 'II',
                    ifelse(line_center_x < center[1] & line_center_y < center[2], 'III', ifelse(line_center_x > center[1] & line_center_y < center[2], 'IV', '')))))


line_data %>% 
  filter(participant == 90, trial_count == 1, is_exterior == 'False') %>% 
  mutate(x_dist = abs(line_center_x - center[1]), y_dist = abs(line_center_y - center[2])) %>% 
  mutate(is_on_axis = ifelse(x_dist < 0.1, 'y_axis', ifelse(y_dist < 0.1, 'x_axis', ''))) %>% 
  mutate(quadrant = ifelse(is_on_axis == '', quadrant, is_on_axis)) %>% 
  group_by(quadrant) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x = quadrant, y = count)) + 
  geom_bar(stat = 'identity') + 
  labs(
    x = 'Quadrant',
    y = 'Number of Lines per Trial'
  ) + 
  theme_bw()

ggsave('lines_per_quadrant.png', height = 1080 / 300, width = 1920 / 300, units = 'in', dpi = 300)

d <- d %>% 
  mutate(x_dist = abs(line_center_x - center[1]), y_dist = abs(line_center_y - center[2])) %>% 
  mutate(is_on_axis = ifelse(x_dist < 0.1, 'y_axis', ifelse(y_dist < 0.1, 'x_axis', ''))) %>% 
  mutate(quadrant = ifelse(is_on_axis == '', quadrant, is_on_axis)) 
  
d %>%   
  filter(!(is.na(quadrant))) %>% 
  group_by(participant, quadrant) %>% 
  summarize(count = n()) %>% 
  group_by(quadrant) %>% 
  summarize(mean_trials = mean(count)) %>% 
  ggplot(aes(x = quadrant, y = mean_trials)) +
  geom_bar(stat = 'identity') +
  labs(
    x = 'Quadrant',
    y = 'Mean Frequency Selected across Participants'
  ) + 
  theme_bw()

ggsave('selection_by_quadrant.png', height = 1080 / 300, width = 1920 / 300, units = 'in', dpi = 300)

d %>% 
  filter(!(is.na(quadrant)), !(is.na(top_x)), !(is.na(accuracy_type))) %>% 
  group_by(participant, quadrant, accuracy_type) %>% 
  summarize(count = n()) %>% 
  group_by(quadrant, accuracy_type) %>% 
  summarize(mean_count = mean(count), se_count = sd(count) / sqrt(N)) %>% 
  ggplot(aes(x = quadrant, y = mean_count, group = accuracy_type)) +
  geom_bar(stat = 'identity', aes(fill = accuracy_type), position = position_dodge(width = .9)) + 
  geom_errorbar(aes(ymin = mean_count - se_count, ymax = mean_count + se_count, group = accuracy_type), width = .5, position = position_dodge(.9)) + 
  labs(
    x = 'Quadrant',
    y = 'Average Across Participants',
    fill = 'Accuracy Type'
  ) + 
  theme_bw() + 
  theme(legend.position = 'bottom')

ggsave('accuracy_by_quadrant.png', height = 1080 / 300, width = 1920 / 300, units = 'in', dpi = 300)

```
  

  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  