---
title: "Grain Boundary Search Task (V1) Preprocessing"
author: "Dave Braun"
date: "12/13/2021"
output: 
  html_document:
    code_folding: hide
    df_print: paged
    theme: flatly
    includes:
      after_body: ../../../../../../html/footer.html
      in_header: ../../../../../../html/favicon.html
    
knit:
  (function(inputFile, encoding) {
      rmarkdown::render(inputFile,
                    encoding = encoding,
                    output_file = 'index.html')})
---

This document was last updated at `r Sys.time()`.

  
```{r include = FALSE}
library(tidyverse)
```

  
  
```{r}
d <- read.csv('../../data/exp1_long_data_raw.csv')
line_data <- read.csv('../../data/exp1_line_data_raw.csv')
head(d)
N <- d %>% 
  group_by(participant) %>% 
  summarize(n()) %>% 
  nrow()
```

**There were `r N` participants in this study.**


### Compute orientation

```{r}
line_data$line_orientation <- ifelse(line_data$top_x == line_data$bottom_x, 'vertical', ifelse(line_data$top_x > line_data$bottom_x, 'forward_slash', ifelse(line_data$top_x < line_data$bottom_x, 'back_slash', 'PROBLEM')))
```


### rt trimming

I want to trim trials where either the initial prompt rt is above 2 sds from participant means or if line selection rt is above 2 sds from p means. Not worrying much about trimming below for now.
I think I'll constrain it such that it has to be above 2 sds *and* 15 s, just so I'm not dropping 2% of data from all ppl.


```{r}
bad_trial_code <- d %>% 
  group_by(participant) %>% 
  summarize(prompt_rt_sec_m = mean(prompt_rt_sec), prompt_rt_sec_sd = sd(prompt_rt_sec), selection_rt_ms_m = mean(selection_rt_ms), selection_rt_ms_sd = sd(selection_rt_ms)) %>% 
  inner_join(d) %>% 
  mutate(bad_trial = ifelse(!is.na(accuracy), 0, ifelse((prompt_rt_sec > (2 * prompt_rt_sec_sd) + prompt_rt_sec_m & prompt_rt_sec > 15) | (selection_rt_ms > (2 * selection_rt_ms_sd) + selection_rt_ms_m & selection_rt_ms > 15000), 1, 0))) %>% 
  filter(bad_trial == 1) %>% 
  group_by(participant, trial_count) %>% 
  summarize(count = n()) %>% 
  mutate(bad_trial = 1) %>% 
  right_join(d) %>% 
  group_by(participant, trial_count) %>% 
  summarize(bad_trial = sum(bad_trial)) %>% 
  mutate(bad_trial = ifelse(is.na(bad_trial), 0, 1)) 

bad_trial_code %>% 
  group_by(participant) %>% 
  summarize(drop_count = sum(bad_trial), count = n()) %>% 
  ggplot(aes(x = drop_count)) +
  geom_histogram(binwidth = 3, color = 'black') + 
  xlab('Percent Trials Dropped') + 
  ylab('Number of Participants')
  
  
```











```{r}
## trim bad trials from thinnest lines data too
thinnest_lines <- read.csv("../../data/thinnest_lines_raw.csv")
thinnest_lines$drop_marker <- paste0(thinnest_lines$participant, '_', thinnest_lines$trial_count)
line_data$drop_marker <- paste0(line_data$participant, '_', line_data$trial_count)
d$drop_marker <- paste0(d$participant, '_', d$trial_count)
bad_trial_code <- unite(bad_trial_code, drop_marker, participant, trial_count, sep = '_')
d <- d %>% 
  inner_join(bad_trial_code)
thinnest_lines <- thinnest_lines %>% 
  inner_join(bad_trial_code)
line_data <- line_data %>% 
  inner_join(bad_trial_code)
```


```{r}
## calculate indecision trial
t <- d %>% 
  group_by(participant, trial_count, selected_or_released) %>% 
  summarize(count = n()) %>% 
  filter(selected_or_released == 'released') %>% 
  mutate(indecision_trial = 1) %>% 
  select(participant, trial_count, indecision_trial)

d <- d %>% 
  left_join(t)

d$indecision_trial <- ifelse(is.na(d$indecision_trial), 0, 1)
```

```{r}
## drop bad trials and save
original_trials <- length(unique(d$participant)) * 100
d <- d[d$bad_trial == 0,]
d <- select(d, -bad_trial)
thinnest_lines <- thinnest_lines[thinnest_lines$bad_trial == 0,]
thinnest_lines <- select(thinnest_lines, -bad_trial)
line_data <- line_data[line_data$bad_trial == 0,]
line_data <- select(line_data, -bad_trial)
final_trials <- d %>% 
  group_by(participant, trial_count) %>% 
  summarize(n()) %>% 
  nrow()
print(paste0('Percent of trials dropped: ', round((1 - (final_trials / original_trials)) * 100, 2), '%'))

```

```{r}
source('../compute_accuracy_columns.r')
d <- compute_accuracy_columns(d, thinnest_lines)
```





```{r}
write.csv(d, '../../data/exp1_long_data.csv', row.names = FALSE)
write.csv(line_data, '../../data/exp1_line_data.csv', row.names = FALSE)
write.csv(thinnest_lines, '../../data/thinnest_lines.csv', row.names = FALSE)
```


















